---
title: "Statistical Inference Project"
author: "Shawn Rich"
output: html_document
---

```{r, echo = FALSE}
# load libraries
library(ggplot2)
```

## Synopsis  
This report will investigate the exponential distribution and compare it with the Central Limit Theorem. This will be done using simulation in R.


## Overview    
For the simulation an exponential distribution having $\lambda = 0.2$ is used. From this 40 random samples are taken and their average calculated. This is repeated 1000 times to create a distribution of means. Recall that for an exponential distribution $\mu = \sigma = 1/\lambda$. For this simulation:

* Let $X_i$ be the outcome for random variable $i$    
* Then $\mu = E[X_i] = 1/0.2 = 5$  
* $Var(X_i) = 5^2 = 25$  
* SE = $\sqrt{\frac{25}{40}} = \frac{5}{2\sqrt{10}} = \frac{5\sqrt{10}}{20}$  

Therefore our **theoretical mean** is **5** and our **theoretical variance** is **25**.

```{r}
#mu = 5
#se = 5*sqrt(10)/20
#mean_distro <- NULL
#for (i in 1 : 1000) mean_distro <- c(mean_distro, (mean(rexp(n=40,rate=0.2)) - mu)/se)


mean_distro <- NULL
for (i in 1 : 1000) mean_distro = c(mean_distro, mean(rexp(40,0.2)))

hist(mean_distro, col="lightgreen")

#data <- data.frame(x = mean_distro, size= factor(rep("Sample Means",1000)))
#g <- ggplot(data, aes(x=x)) + geom_histogram(binwidth=.3, colour = "black", #fill = "lightgreen", aes(y = ..density..)) 
#g <- g + stat_function(fun = dnorm, size = 2)
#g + scale_x_continuous(limits=c(2,3,4,5,6,7,8), breaks=1:7)
#g + facet_grid(. ~ size)
```


The mean is `r round(mean(mean_distro),2)` and variance is `r round(var(mean_distro),2)`

## Sample Mean versus Theoretical Mean  


## Sample Variance versus Theoretical Variance


## Distribution
