---
title: "ToothGrowth Data Analysis"
author: "Shawn Rich"
date: "09/26/2015"
output: html_document
---

```{r, echo = FALSE}
# load libraries
library(ggplot2)
library(gridExtra)
library(datasets)
```

## Synopsis

The goal of this report is to provide analysis of the ToothGrowth dataset from the R datasets package. 

## Code Book

From documentation of the ToothGrowth dataset we know it contains data from a study of the effect of vitamin C on tooth growth in 60 guinea pigs. The data consist of 60 observations on three variables. The variables are:

**len** (numeric) - Tooth Length    
**supp** (factor) - Type of vitamin supplement. Orange juice *OJ* or vitamin C supplement *VC*  
**dose** (numeric) - Dosage in milligrams per day.  

## Summary of Data

Load the data and perform some basic exploration to see if it matches what we know from the documentation.


```{r}
attach(ToothGrowth)
# number of rows in dataset
nrow(ToothGrowth)
# structure of data
str(ToothGrowth)
# cross tab of supp vs dose
table(ToothGrowth$supp,ToothGrowth$dose)
```

The table shows that there are 10 guinea pigs in each of the six dosage/supplement groups. It will be assumed that there is no overlap in the groups and therefore independent.


## Analysis

For our analysis we first split the data base on supplement.

```{r}
OJ <- data.frame(len = ToothGrowth$len[supp=="OJ"],
                 dose = ToothGrowth$dose[supp=="OJ"])
VC <- data.frame(len = ToothGrowth$len[supp=="VC"],
                 dose = ToothGrowth$dose[supp=="VC"])
```

<!--
Now in the second portion of the class, we're going to analyze the ToothGrowth data in the R datasets package. 

    Load the ToothGrowth data and perform some basic exploratory data analyses 
    Provide a basic summary of the data.
    Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering)
    State your conclusions and the assumptions needed for your conclusions. 

Some criteria that you will be evaluated on

    Did you  perform an exploratory data analysis of at least a single plot or table highlighting basic features of the data?
    Did the student perform some relevant confidence intervals and/or tests?
    Were the results of the tests and/or intervals interpreted in the context of the problem correctly? 
    Did the student describe the assumptions needed for their conclusions?
-->
