---
title: "Analysis of ToothGrowth Data using T Test"
author: "Shawn Rich"
date: "09/26/2015"
output: pdf_document
---

## Synopsis

The goal of this report is to provide analysis of the ToothGrowth dataset from the R datasets package. Specifically we will use R t.test function to investigate whether increased amounts of vitamin C supplement increase tooth growth in guinea pigs.

## Summary of Data

First we load the data and perform some basic exploration.

```{r}
library(datasets)
attach(ToothGrowth)
# structure of data
str(ToothGrowth)

```

The variables are:  
**len** (numeric) - Tooth Length    
**supp** (factor) - Type of vitamin supplement. Orange juice *OJ* or vitamin C supplement *VC*  
**dose** (numeric) - Dosage in milligrams per day.  

The table below shows the samples sizes. There are 10 guinea pigs in each of the six dosage/supplement groups which accounts for all 60 observations.

```{r}
# cross tab of supp vs dose
table(ToothGrowth$supp,ToothGrowth$dose)
```


## Assumptions

1. It will be assumed that there is no overlap in the sample groups and are therefore independent. Base on this we will use *paired = FALSE* in our t.test.  
2. We also assume that the variances among the groups are not equal and will use *var.equal = FALSE* in our t.test.


## Analysis

For our analysis we first split the data based on supplement type and dosage size. For the analysis we will only compare the two dosage groups of 0.5mg/day and 2.0mg/day.

```{r}
OJ05 <- data.frame(len = ToothGrowth$len[supp=="OJ" & dose == 0.5])
OJ20 <- data.frame(len = ToothGrowth$len[supp=="OJ" & dose == 2.0])
VC05 <- data.frame(len = ToothGrowth$len[supp=="VC" & dose == 0.5])
VC20 <- data.frame(len = ToothGrowth$len[supp=="VC" & dose == 2.0])
```


Using these data vectors of tooth length base on dosage and supplement type we examine effects of vitamin C on tooth length. For this we conduct t.test for each form of vitamin C comparing the dose 0.5mg/day with that of 2.0mg/day. We wish to determine if higher doses of vitamin C increase tooth growth. Our null hypothesis and alternative hypothesis are as follows:

$H_{o} = \mu_{1} = \mu_{2}$ or $\mu_{1} - \mu_{2} = 0$  
$H_{a} = \mu_{1} < \mu_{2}$ or $\mu_{1} - \mu_{2} < 0$

**Orange Juice**  
Below is output of t.test for effect of OJ on tooth length at dose 0.5 mg/day and 2.0 mg/day.

```{r}
t.test(OJ05, OJ20, alternate=c("twosided","less"), 
       paired = FALSE, var.equal = FALSE)
```

The 95% confidence interval lies completely below 0 and therefore unlikely that $\mu_{1} - \mu_{2} = 0$. So, we can reject null hypothesis in favor of the alternate hypothesis. Increase levels of vitamin C from orange juice would appear to increase tooth growth.

**Vitamin C Supplement**  
Below is output of t.test for effect of VC on tooth length at dose 0.5 mg/day and 2.0 mg/day.

```{r}
t.test(VC05, VC20, alternate=c("twosided","less"), 
       paired = FALSE, var.equal = FALSE)
```

Again the 95% confidence interval lies completely below 0 and therefore unlikely that $\mu_{1} - \mu_{2} = 0$. So, we can reject null hypothesis in favor of the alternate hypothesis. Increase levels of vitamin C supplement would appear to increase tooth growth.

**OJ vs VC**  
Lastly, is there a difference between the supplements? For this we compare OJ20 to VC20. Our null hypothesis and alternative hypothesis remains as before.

```{r}
t.test(OJ20, VC20, alternate=c("twosided","less"), 
       paired = FALSE, var.equal = FALSE)
```

Here the 95% confidence interval contains 0. In fact, zero appears to be quite nearly in the center of the interval. The p-value is also large. Based on this we accept the null hypothesis and conclude that there is no difference in type of vitamin C supplement on tooth growth. 
