# Regmods Week 4 Problem Set


## Generalized Linear Models

1. True or false, generalized linear models transform the observed outcome.

```{r}

```


2. True or false, the interpretation of the coefficients in a GLM are on the scale of the link function. (Discuss.)

```{r}

```


3. True or false, the generalized linear model assumes an exponential family for the outcome. (Discuss.)

```{r}

```


4. True or false, GLM estimates are obtained by maximizing the likelihood. (Discuss.)

```{r}

```


5. True or false, some GLM distributions impose restrictions on the relationship between the mean and the variance. (Discuss.)

```{r}

```


## Binary GLMs

1. Load the dataset Seatbelts as part of the datasets package via data(Seatbelts). Use as.data.frame to convert the object to a dataframe. Create a new outcome variable for whether or not greater than 119 drivers were killed that month. Fit a logistic regression GLM with this variable as the outcome and kms, PetrolPrice and law as predictors. Interpret your parameters. 

```{r}

```

2. Fit a binomial model with DriversKilled as the outcome and drivers as the total count with kms , PetrolPrice and law as predictors, interpret your results.

```{r}

```

3. Refer to Question 1. Use the anova function to compare models with just law, law and PetrolPrice and all three predictors. 

```{r}

```


## Count Data

1. Load the dataset Seatbelts as part of the datasets package via data(Seatbelts). Use as.data.frame to convert the object to a dataframe. Fit a Poisson regression GLM with UKDriversKilled as the outcome and kms, PetrolPrice and law as predictors. Interpret your results.

2. Refer to question 1. Fit a linear model with the log of drivers killed as the outcome. Interpret your results.

3. Refer to question 1. Fit your Poisson log-linear model with drivers as a log offset (to consider the proportion of drivers killed of those killed or seriously injured.)

4. Refer to Question 1. Use the anova function to compare models with just law, law and PetrolPrice and all three predictors.

