---
title: "Statistical Inference Homework 2"
output:
  html_document:
    theme: united
    toc: yes
  pdf_document:
    highlight: zenburn
    toc: yes
---

## Question 1  
The probability that a manuscript gets accepted to a journal is 12% (say). However, given that a revision is asked for, the probability that it gets accepted is 90%. Is it possible that the probability that a manuscript has a revision asked for is 20%?

1. Yeah, that's totally possible.
1. No, it's not possible.
1. It's not possible to answer this question.

**Answer**  
A = manuscript accepted  
B = revision asked for  
$P(A) = 0.12$  
$P(B) = 0.20$ (is this possible?)  
$P(A|B) = 0.90$  

$P(A|B) = P(A \cap B) / P(B)$  
$0.90 = P(A \cap B) / 0.20$  
$0.18 = P(A \cap B)$  

The probability of $P(A \cap B)$ cannot be greater than $P(A)$. Therefore, it is not possible that the probability that a manuscript has a revision asked for is 20%.


## Question 2  
Suppose that the number of web hits to a particular site are approximately normally distributed with a mean of 100 hits per day and a standard deviation of 10 hits per day. What's the probability that a given day has fewer than 93 hits per day expressed as a percentage to the nearest percentage point?

1. 76%
1. 24%
1. 47%
1. 94%

**Answer**  

```{r}
pnorm(93, mean=100, sd=10, lower.tail=TRUE)
```

We used **lower.tail=TRUE** because we want the $P[X \le x]$, where x = 93 in this problem. Use **lower.tail=FALSE** for $P[X \gt x]$


## Question 3  
Suppose 5% of housing projects have issues with asbestos. The sensitivity of a test for asbestos is 93% and the specificity is 88%. What is the probability that a housing project has no asbestos given a negative test expressed as a percentage to the nearest percentage point?

1. 0%
1. 5%
1. 10%
1. 20%
1. 50%
1. 100%

**Answer**  
D is event has abestos  
+ test is positive  
- test is negative  
$P(D) = 0.05$  
Sensitivety of test = 0.93 = $P(+|D)$  
Specificity of test = 0.88 = $P(-|~D)$  

What is $P(~D|-)$?

From **Bayes' Rule:**  
$P(B|A) = \frac{P(A|B) * P(B)}{P(A|B) * P(B) + P(A|~B) * P(~B)}$  

we get:  
$P(~D|-) = \frac{P(-|~D) * P(~D)}{P(-|~D) * P(~D) + P(-|D) * P(D)}$  

We have everything except $P(~D)$ and $P(-|D)$

From the swirl lession recall that:  
$P(~D) = 1 - P(D) = 1 - 0.05 = 0.95$  
and  
$P(-|D) = 1 - P(+|D) = 1 - 0.93 = 0.07$  
We now have everything and just plug in the values into Bayes' rule:  
```{r}
(0.88*0.95) / (0.88*0.95 + 0.07*0.05)
```

$P(~D|-)$ is 100%


## Question 4  
Suppose that the number of web hits to a particular site are approximately normally distributed with a mean of 100 hits per day and a standard deviation of 10 hits per day.  

What number of web hits per day represents the number so that only 5% of days have more hits? Express your answer to 3 decimal places.  

```{r}
round(qnorm(0.05, mean=100, sd=10, lower.tail=FALSE),3)
```


## Question 5  
Suppose that the number of web hits to a particular site are approximately normally distributed with a mean of 100 hits per day and a standard deviation of 10 hits per day.

Imagine taking a random sample of 50 days. What number of web hits would be the point so that only 5% of averages of 50 days of web traffic have more hits? Express your answer to 3 decimal places. 

HINT: Let $\bar{X}$ be the average number of hits per day for 50 randomly sampled days. X is N(100,10^2/50). Therefore sd = $10 / \sqrt[2]{50}$

```{r}
round(qnorm(.05, mean = 100, sd = 10 / sqrt(50), lower.tail = FALSE), 3)
```


## Question 6  
You don't believe that your friend can discern good wine from cheap. Assuming that you're right, in a blind test where you randomize 6 paired varieties (Merlot, Chianti, ...) of cheap and expensive wines

What is the chance that she gets 5 or 6 right expressed as a percentage to one decimal place?

${6 \choose 5} * .5^5(1 - .5)^1 + {6 \choose 6} * .5^6(1 - .5)^0 \approx 0.109$  
$0.109 * 100 = 10.9$

```{r}
round((choose(6, 5) * 0.5^6 + choose(6, 6) * 0.5^6)*100,1)
```

**or**

```{r}
round(pbinom(4, size=6, prob=0.5,lower.tail= FALSE)*100,1)
```

## Question 7  
Consider a uniform distribution. If we were to sample 100 draws from a a uniform distribution (which has mean 0.5, and variance 1/12) and take their mean, $\bar{X}$

What is the approximate probability of getting as large as 0.51 or larger expressed to 3 decimal places?


## Question 8  
If you roll ten standard dice, take their average, then repeat this process over and over and construct a histogram, what would it be centered at?


## Question 9  
If you roll ten standard dice, take their average, then repeat this process over and over and construct a histogram, what would be its variance expressed to 3 decimal places?


## Question 10  
The number of web hits to a site is Poisson with mean 16.5 per day. What is the probability of getting 20 or fewer in 2 days expressed as a percentage to one decimal place?


