---
title: "Statistical Inference Homework 3"
output:
  html_document:
    theme: united
    toc: no
  pdf_document:
    toc: no
---

## Question 1  
Load the data set mtcars in the datasets R package. Calculate a 95% confidence interval to the nearest MPG for the variable mpg.  
- What is the lower endpoint of the interval?  **18**  
- What is the upper endpoint of the interval?  **22**  

```{r}
# load mtcars dataset
library(datasets)
data(mtcars)
# attach so you don't have to use $ notation as in: mtcars$mpg
attach(mtcars)
# extra stuff
head(mtcars)
mean(mpg)
sd(mpg)
# answer
round(t.test(mpg)$conf.int)
```

## Question 2  
Suppose that standard deviation of 9 paired differences is 1. What value would the average difference have to be so that the lower endpoint of a 95% students t confidence interval touches zero?  
- Give the number here to two decimal places  **0.77**  

n = 9  
sd = 1  
ybar +- t.975 * 1/3 where 1/3 = 1/sqrt(n) = 1/sqrt(9)
$\bar{y} +- t.975 1/3 = 0$
```{r}
round(qt(0.975, 8)/3,2)
```


## Question 3  
An independent group Student's T interval is used instead of a paired T interval when:  
- The observations are paired between the groups.  
**X** The observations between the groups are naturally assumed to be statistically independent  
- As long as you do it correctly, either is fine.  
- More details are needed to answer this question  


## Question 4  
Consider the mtcars dataset. Construct a 95% T interval for MPG comparing 4 to 6 cylinder cars (subtracting in the order of 4 - 6) assume a constant variance.  
- What is the lower endpoint of the interval to 1 decimal place?  **3.15**  
- What is the upper endpoint of the interval to 1 decimal place?  **10.69**  

Answer: 
```{r}
head(mtcars)
c4 <- mtcars$mpg[mtcars$cyl == 4]
c6 <- mtcars$mpg[mtcars$cyl == 6]
t.test(c4, c6, var.equal = TRUE)

# alternate
round(as.vector(t.test(c4, c6, var.equal = TRUE)$conf.int),1)
```
## Question 5  
If someone put a gun to your head and said "Your confidence interval must contain what it's estimating or I'll pull the trigger", what would be the smart thing to do?  
**X** Make your interval as wide as possible  
- Make your interval as small as possible  
- Call the authorities  


## Question 6  
Refer back to comparing MPG for 4 versus 6 cylinders. What do you conclude?  
- The interval is above zero, suggesting 6 is better than 4 in the terms of MPG  
**X** The interval is above zero, suggesting 4 is better than 6 in the terms of MPG  
- The interval does not tell you anything about the hypothesis test; you have to do the test.  
- The interval contains 0 suggesting no difference.  


## Question 7  
Suppose that 18 obese subjects were randomized, 9 each, to a new diet pill and a placebo. Subjects' body mass indices (BMIs) were measured at a baseline and again after having received the treatment or placebo for four weeks. The average difference from follow-up to the baseline (followup - baseline) was 3 kg/m2 for the treated group and 1 kg/m2 for the placebo group. The corresponding standard deviations of the differences was 1.5 kg/m2 for the treatment group and 1.8 kg/m2 for the placebo group. The study aims to answer whether the change in BMI over the four week period appear to differ between the treated and placebo groups.  
- What is the pooled variance estimate? (to 2 decimal places)  

Answer:  
Since the the two groups are same size, the pooled variance is the average of the two. We are given standard deviations, so we need to convert to variance.  
```{r}
round((1.5^2 + 1.8^2)/2,2)
```

A general solution whould be:

```{r}
n1 <- n2 <- 9
x1 <- -3  ##treated
x2 <- 1  ##placebo
s1 <- 1.5  ##treated
s2 <- 1.8  ##placebo
spsq <- ( (n1 - 1) * s1^2 + (n2 - 1) * s2^2) / (n1 + n2 - 2)
round(spsq,2)
```


## Question 8 (hold out from previous version of class)   
For Binomial data the maximum likelihood estimate for the probability of a success is  
**X** The proportion of successes  
- The proportion of failures  
- A shrunken version of the proportion of successes  
- A shrunken version of the proportion of failures  


## Question 9 (hold out from previous version of class)  
Bayesian inference requires  
- A type I error rate  
- Setting your confidence level  
**X** Assigning a prior probability distribution  
- Evaluating frequency error rates  
