---
output: pdf_document
fontsize: 12pt
geometry: margin=0.75in
---


**The Birthday Problem**

In a group of *n* people, what is the probability that at least two will have the same birthday? Below is a graph showing group sizes up to 90 and the corresponding probability that at least two people in the group will have the same birthday.

Let B = the probability that at least two people of the same birthday. Then:

$P(B) = 1 - \frac{365}{365} * \frac{364}{365} * \frac{363}{365} * ...\frac{365 - (n-1)}{365}$

In the above formula we calculate the probability that no one has the same birthday and subtract this from 1. This approach make the calculation easier. The **R** code below implements this formula in the *sapply* function.


```{r echo=FALSE}
library(ggplot2)
```

```{r fig.width=6, fig.height = 4}
# calculate the probability of two people having same birthday
dat <- data.frame(group_size = c(1:90))
dat$probability <- sapply(dat$group_size,function(x) 
  1 - prod(365:(365 - (x -1))) / 365^x)

# plot the results
ggplot(data=dat, aes(x = group_size, y=probability)) +
  geom_line(colour="black") +
  coord_cartesian(xlim = c(0,90)) + scale_x_continuous(breaks=seq(0,90,10))  +
  coord_cartesian(ylim = c(0.00,1.00)) + 
  scale_y_continuous(breaks=seq(0.00,1.00,0.10)) + 
  geom_vline(aes(xintercept=23), linetype="dashed")
```

At a group size of only 23, the probability that at least two have the same birthday is 50%. 
