---
title: "Birthday Problem"
author: "Shawn Rich"
date: "October 19, 2015"
output: pdf_document
fontsize: 11pt
geometry: margin=0.75in
---

**Overview**

In a group of people, what is the probability of at least two people having the same birthday? The plot below shows the probabiltity of group sizes up to 90. 

```{r echo=FALSE}
library(ggplot2)
```

```{r fig.width=5, fig.height = 3.5}
# calculate the probability of two people having same birthday
# in group sizes up to 90 people
dat <- data.frame(group_size = c(1:90))
dat$probability <- sapply(dat$group_size,function(x) 
  1 - prod(365:(365 - (x -1))) / 365^x)

# plot the results
ggplot(data=dat, aes(x = group_size, y=probability)) +
  geom_line(colour="black") +
  coord_cartesian(xlim = c(0,90)) + scale_x_continuous(breaks=seq(0,90,10))  +
  coord_cartesian(ylim = c(0.00,1.00)) + 
  scale_y_continuous(breaks=seq(0.00,1.00,0.10))
```


* What does the group size need to be in order to have a 50% chance that two people will have the same birthday?
* What about a 70% chance?
* Group sizes over what size will give us a nearly 100% chance of two people having same birthday?