---
title: "Machine Learning Project"
author: "Shawn Rich"
date: "May 7, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r warning=FALSE, echo=FALSE, error=FALSE, message=FALSE}
# initializations
rm(list=ls())

# set working directory
setwd("~/Documents/courses/course_notes/machine_learning/project")

# load libraries
library(caret)

```

## Load Train and Test Data

```{r warning=FALSE, echo=FALSE, error=FALSE, message=FALSE}
trainingData <- read.table("./data/pml-training.csv", sep=",", na.strings = c("NA","#DIV/0!"),header = TRUE)
testingData <- read.table("./data/pml-testing.csv", sep=",", na.strings = c("NA","#DIV/0!"),header = TRUE)
```

## Cleanup Data

The first two columns "X" and "user_name" are removed as they are irrelevant as predictors. The data contain many NA values. These are set to zero. Next the caret function nearZeroVar to identify near zero-variance variables that we would want to eliminate as predictors in our model.

```{r warning=FALSE, echo=TRUE, error=FALSE, message=FALSE}
trainingData <- trainingData[,-c(1,2)]
trainingData[is.na(trainingData)] <- 0

dim(trainingData)
nzv <- nearZeroVar(trainingData)
trainingData <- trainingData[,-nzv]
dim(trainingData)
```
nzv[nzv$zeroVar,]
head(trainingData)
summary(training$X)
head(training$X)
tail(training$X)
colnames(trainingData)
colnames(testingData)
summary(trainingData$classe)
summary(testingData$problem_id)
head(testingData$problem_id)
head(pml_training[which(pml_training$new_window=="yes"),])
summary(trainingData$user_name)

