---
title: "Regression Models Exercises - Week 2"
author: "Shawn Rich"
date: "January 12, 2016"
output: html_document
---

```{r echo=FALSE, results="hide", warning=FALSE, message=FALSE}
library(UsingR)
data("galton")
data("father.son")
```


## Statistical linear regression models

1. Fit a linear regression model to the father.son dataset with the father as the predictor and the son as the outcome. Give a p-value for the slope coefficient and perform the relevant hypothesis test.

Recall: $SH = \beta_0 + \beta_1 * FH$

* $H_0: \beta_1 = 0$
* $H_a: \beta_1 \neq 0$

From the output of summary(fit)$coef we see a P values of 1.2e-69 and t value of 19. We this we reject the null hypothesis and accept the alternate and conclude that the father's height does affect the son's height.

```{r}
x <- father.son$fheight
y <- father.son$sheight
fit <- lm(y ~ x)
summary(fit)$coef
```

2. Refer to question 1. Interpret both parameters. Recenter for the intercept if necessary.

```{r}

```

3. Refer to question 1. Predict the son’s height if the father’s height is 80 inches. Would you recommend this prediction? Why or why not?

```{r}

```

4. Load the mtcars dataset. Fit a linear regression with miles per gallon as the outcome and horsepower as the predictor. Interpret your coefficients, recenter for the intercept if necessary. 

```{r}

```

5. Refer to question 4. Overlay the fit onto a scatterplot.


```{r}

```

6. Refer to question 4. Test the hypothesis of no linear relationship between horsepower and miles per gallon. 


```{r}

```

7. Refer to question 4. Predict the miles per gallon for a horsepower of 111. 


```{r}

```


