---
title: "Relationship of Variables and Miles Per Gallon"
author: "Shawn Rich"
date: "January 26, 2016"
output: html_document
---

```{r warning=FALSE, echo=FALSE, error=FALSE, message=FALSE}
# load libraries
library(UsingR)
data("mtcars")
```

## Synopsis

This report, using the the mtcars dataset, examines the relationship among the variables in the dataset and miles per gallon (MPG) as outcome. Specifically addressed is  whether the an automatic or manual transmission results in better gas milage.

## Executive Summary


## Data Exploration

Before building models and analysis we need an understanding of the data. Appendix A list the variables in the dataset with their descriptions and the first rows of the dataset. Next we see if a null hypothesis is true. Specifically is the difference of the mean mpg of automatic vs manual equal to zero? Visually we can see this in the boxplot. The t-test, with p-value of 0.0014, gives us more confidence that we can reject the null in favor the alternate hypothesis. There is a diffence in mile per gallon. The next step then is to develop a model of these data. 


## Data Model

First we look at the correlation among all variables against *mpg* to see which should be included in our model. From the correlation matrix we that *cyl*, *disp*, *hp*, and *wt* are highly correlated with mgp. There is also a high correlation of `r cor(mtcars$disp, mtcars$cyl)` between *disp* and *cyl*, which mean we may not want to inlcude both. This correlation make intuitive sense, because more cyclinders the more overall displacement an engine would have.


## Appendix A - Detail Data Exploration

The MTCARS dataset variables and their descriptions.

~~~~~
mpg         Miles/(US) gallon  
cyl         Number of cylinders  
disp        Displacement (cu.in.)  
hp          Gross horsepower  
drat        Rear axle ratio  
wt          Weight (1000 lbs)  
qsec        1/4 mile time  
vs          Cylinder alignment (0 - V; 1 - straight)
am          Transmission type (0 = automatic, 1 = manual)  
gear        Number of forward gears  
carb        Number of carburetors  
~~~~~

The first 10 rows of the dataset to get a overall view of the data.

```{r echo=FALSE}
head(mtcars)
```

## Apendix B - Detail Data Exploration

**Boxplot MPG vs Tranmission Type**

```{r fig.width=5, fig.height = 6}
boxplot(mpg ~ am, data=mtcars, xlab="Transmission Type",
ylab="Miles per Gallon",
main="Automatic versus Manual Transmission MPG")
```

**T-Test**

```{r}
mpg.auto <- mtcars[mtcars$am == 0,]$mpg
mpg.manual <- mtcars[mtcars$am == 1,]$mpg

t.test(mpg.auto, mpg.manual)
```


## Appendix C - Data Model Selection

**Correlation of Variables**

Variables *cyl*, *disp*, *hp*, and *wt* are highly correlated with mgp. The high correlation between *disp* and *cyl* indicates that those two should be examined and determine if both should be included. 

```{r}
cor(mtcars)[1,]
cor(mtcars$disp, mtcars$cyl)
```

**The Models**

Five nested models are created below. These are:

Model 1: mpg ~ am
Model 2: mpg ~ am + hp
Model 3: mpg ~ am + hp + wt
Model 4: mpg ~ am + hp + wt + cyl
Model 5: mpg ~ am + hp + wt + disp

Models 4 and 5 are used to see if adding either *cyl* or *disp* is necessary. ANOVA test are done to compare the models and help make our model selection.

```{r}
model1 <- lm(mpg ~ am, data = mtcars)
model2 <- lm(mpg ~ am + hp, data = mtcars)
model3 <- lm(mpg ~ am + hp + wt, data = mtcars)
model4 <- lm(mpg ~ am + hp + wt + cyl, data = mtcars)
model5 <- lm(mpg ~ am + hp + wt + disp, data = mtcars)
anova(model1, model2, model3, model4)
anova(model1, model2, model3, model5)
```

From both ANOVA test the p-values indicate that adding *hp* and *wt* is necessary. Both having p-values less than 0.05. However, both *cyl* and *disp* have high p-values (> 0.05) indicating that neither is necessary in the models. The final model then is model 3 inlcuding mpg *am*, *hp*, and *wt* predictors and *mpg* as outcome.

```{r}
summary(model3)
```