---
title: "Regression Models Exercises"
author: "Shawn Rich"
date: "November 13, 2015"
output: html_document
---
```{r echo=FALSE, results="hide", warning=FALSE, message=FALSE}
library(UsingR)
```

## Introduction

1. Consider the dataset given by x=c(0.725,0.429,-0.372 ,0.863). What value of mu minimizes sum((x - mu)^2)?

$X_bar$ is the value of mu that minimizes for a dataset.

```{r}
x <- c(0.725,0.429,-0.372 ,0.863)
mean(x)
```

2. Reconsider the previous question. Suppose that weights were given, w = c(2, 2, 1, 1) so that we wanted to minimize sum(w * (x - mu) ^ 2) for mu. What value would we obtain? 

```{r}
x <- c(0.725,0.429,-0.372 ,0.863)
w <- c(2, 2, 1, 1)
xw <- x * w
mean(xw)
```

3. Take the Galton and obtain the regression through the origin slope estimate where the centered parental height is the outcome and the child’s height is the predictor.

```{r}
data(galton)
y <- galton$parent
x <- galton$child
yc <- y - mean(y)
xc <- x - mean(x)
sum(yc * xc) / sum(xc^2)
```

We get the same result with lm.

```{r}
lm(formula = yc ~ xc - 1)
```

We get same slope coefficient with y as the outcome and x as the predictor.

```{r}
lm(formula = y ~ x)
```

## Notation

1. Take the Galton dataset and find the mean, standard deviation and correlation between the parental and child heights.

```{r}
data(galton)
summary(galton)
cor(galton)
```

2. Center the parent and child variables and verify that the centered variable means are 0. 

```{r}

```

3. Rescale the parent and child variables and verify that the scaled variable standard deviations are 1. 

```{r}

```

4. Normalize the parental and child heights. Verify that the normalized variables have mean 0 and standard deviation 1 and take the correlation between them. 

```{r}

```
