---
title: "Analysis of Miles Per Gallon for Automatic vs Manual Transmissions"
date: "January 31, 2016"
output: pdf_document
---

```{r warning=FALSE, echo=FALSE, error=FALSE, message=FALSE}
# load libraries
library(UsingR)
data("mtcars")
```

## Executive Summary

This report, using the the mtcars data-set, examines the relationship among the variables in the data-set and miles per gallon (*mpg*) as outcome. Two main points addressed in this report are (1) whether an automatic or manual transmission results in better gas mileage and (2) what is the difference in *mpg* between automatic and manual transmissions.

Part of the data exploration was to determine if there is a difference in mean *mpg* for automatic and manual transmissions. A t-test shows that a null hypothesis of means being equal should be rejected. Further analysis was then justified. 

Analysis started by finding other variables that are highly correlated with *mpg*. These are *cyl*, *disp*, *hp*, and *wt*. Variables *cyl* and *disp* were also found to be highly correlated with each other. Five nested models were created starting with *mpg ~ am* and then adding others individually. ANOVA was used to compare these models. During this comparison it was found the neither *cyl* or *disp* were necessary. The final model was *mpg ~ am + hp + wt*. This models shows that on average there is a **2.084** miles per gallon increase for a manual transmission over an automatic. The $R^2$ value shows the model explains 84% of the variation. Residual plot did not show aspects of poor model fit.


## Data Exploration

Before building models and analysis we need an understanding of the data. Below are listed the variables in the data-set with their descriptions and the first rows of the data-set. Next we see if a null hypothesis is true. Specifically is the difference of the mean mpg of automatic vs manual equal to zero? The t-test, with p-value of 0.0014, gives us confidence that we can reject the null in favour the alternate hypothesis that there is a difference in miles per gallon. The next step then is to develop a model of these data. 

**The MTCARS data-set variables and their descriptions.**

~~~~~
mpg         Miles/(US) gallon  
cyl         Number of cylinders  
disp        Displacement (cu.in.)  
hp          Gross horsepower  
drat        Rear axle ratio  
wt          Weight (1000 lbs)  
qsec        1/4 mile time  
vs          Cylinder alignment (0 - V; 1 - straight)
am          Transmission type (0 = automatic, 1 = manual)  
gear        Number of forward gears  
carb        Number of carburetors  
~~~~~

The first 10 rows of the data-set to get a overall view of the data.

```{r echo=FALSE}
head(mtcars)
```

**T-Test**

$H_0: \mu_a - \mu_m  = 0$

$H_a: \mu_a - \mu_m  \neq 0$

where:  
  $\mu_a$ = average mpg of automatic  
  $\mu_m$ = average mpg of manual  

With with p-value of 0.0014 we reject $H_0$ in favour of $H_a$.

```{r}
mpg.auto <- mtcars[mtcars$am == 0,]$mpg
mpg.manual <- mtcars[mtcars$am == 1,]$mpg

t.test(mpg.auto, mpg.manual)
```


## Data Model

First we look at the correlation among all variables against *mpg* to see which should be included in our model. From the correlation matrix we that *cyl*, *disp*, *hp*, and *wt* are highly correlated with mpg. There is also a high correlation of `r cor(mtcars$disp, mtcars$cyl)` between *disp* and *cyl*, which means we may not want to include both. 

```{r}
cor(mtcars)[1,]
cor(mtcars$cyl,mtcars$disp)
```

**The Models**

Five nested models are created below, **am** being included in all. These are:

Model 1: mpg ~ am  
Model 2: mpg ~ am + hp  
Model 3: mpg ~ am + hp + wt  
Model 4: mpg ~ am + hp + wt + cyl  
Model 5: mpg ~ am + hp + wt + disp  

Models 4 and 5 are used to see if adding either *cyl* or *disp* is necessary. ANOVA test are done to compare the models and help make our model selection.

```{r}
model1 <- lm(mpg ~ am, data = mtcars)
model2 <- lm(mpg ~ am + hp, data = mtcars)
model3 <- lm(mpg ~ am + hp + wt, data = mtcars)
model4 <- lm(mpg ~ am + hp + wt + cyl, data = mtcars)
model5 <- lm(mpg ~ am + hp + wt + disp, data = mtcars)
anova(model1, model2, model3, model4)
anova(model1, model2, model3, model5)
```

From both ANOVA test the p-values indicate that adding *hp* and *wt* is necessary. Both having p-values less than 0.05. However, both *cyl* and *disp* have high p-values (> 0.05) indicating that neither is necessary in the models. The final model then is model 3 including mpg *am*, *hp*, and *wt* as predictors and *mpg* as outcome.

Before looking at the summary of model3, we examine residuals to diagnose normality of the errors. In the Normal Q-Q plot below we see that the residuals fall roughly on the line indicating a fairly good model fit.

```{r, fig.width=9, fig.height = 6 }
plot(model3, which=2)
```

```{r}
summary(model3)
```

The summary shows that on average there is a **2.084** mile per gallon increase for a manual transmission over an automatic. The $R^2$ value show the model explains **84%** of the variation. 