---
title: "Relationship of Variables and Miles Per Gallon"
author: "Shawn Rich"
date: "January 26, 2016"
output: pdf_document
---

```{r warning=FALSE, echo=FALSE, error=FALSE, message=FALSE}
# load libraries
library(UsingR)
data("mtcars")
```

## Synopsis

In this report I will use the mtcars dataset to examine the relationship among the variables in the data and miles per gallon (MPG) as outcome. Specifically I will address whether the an automatic or manual transmission results in better gas milage for the car. To this end I will quantify the MPG difference between automatic and manual transmissions.

## Executive Summary


## Data Exploration

Before building models and analysis we need an understanding of the data. Appendix A list the variables in the dataset with their descriptions and the first rows of the dataset. Next we see if a null hypothesis is true. Specifically is the difference of the mean mpg of automatic vs manual equal zero? Visually we can see this in the boxplot. The t-test, with p-value of 0.0014, gives us more confidence that we can reject the null in favor the alternate hypothesis. There is a diffence in mile per gallon. 


## Data Analysis



## Appendix A - Detail Data Exploration

The MTCARS variables and their descriptions.

~~~~~
mpg         Miles/(US) gallon  
cyl         Number of cylinders  
disp        Displacement (cu.in.)  
hp          Gross horsepower  
drat        Rear axle ratio  
wt          Weight (1000 lbs)  
qsec        1/4 mile time  
vs          Cylinder alignment (0 - V; 1 - straight)
am          Transmission type (0 = automatic, 1 = manual)  
gear        Number of forward gears  
carb        Number of carburetors  
~~~~~

The first 10 rows of the dataset to get a overall view of the data.

```{r echo=FALSE}
head(mtcars)
```

## Apendix B - Detail Data Exploration

**Data Cleanup**

```{r}
# convert am 0,1 values to factor variables auto, manual respectively
mtcars$am <- as.factor(mtcars$am)
levels(mtcars$am) <-c("auto", "manual")
```

**Simple Model Fit**

The simple_model uses just mpg as outcome and am as predictor.

```{r}
simple_model <- lm(mpg ~ am, data = mtcars)
summary(simple_model)$coef
confint(simple_model)
```

```{r fig.width=5, fig.height = 6}
boxplot(mpg ~ am, data=mtcars, xlab="Transmission Type",
ylab="Miles per Gallon",
main="Automatic versus Manual Transmission MPG")
```

```{r}
mpg.auto <- mtcars[mtcars$am == "auto",]$mpg
mpg.manual <- mtcars[mtcars$am == "manual",]$mpg

t.test(mpg.auto, mpg.manual)
```