---
title: "Relationship of Variables and Miles Per Gallon"
author: "Shawn Rich"
date: "January 26, 2016"
output: html_document
---

```{r warning=FALSE, echo=FALSE, error=FALSE, message=FALSE}
# load libraries
library(UsingR)
data("mtcars")
```

## Summary

In this report I will use the mtcars dataset to examine the relationship among the variables in the data and miles per gallon (MPG) as outcome. Specifically I will address whether the an automatic or manual transmission results in better gas milage for the car. To this end I will quantify the MPG difference between automatic and manual transmissions.

## Data Exploration

Before building models and analysis we need an understanding of the data. Appendix A list the variables in the dataset with their descriptions and the first rows of the dataset. We now know what the variables mean and what type of data they contain.

```{r}
head(mtcars)
```

## Appendix A - MTCARS Dataset Description

The MTCARS variables and their descriptions.

~~~~~
mpg         Miles/(US) gallon  
cyl         Number of cylinders  
disp        Displacement (cu.in.)  
hp          Gross horsepower  
drat        Rear axle ratio  
wt          Weight (1000 lbs)  
qsec        1/4 mile time  
vs          Cylinder alignment (0 - V; 1 - straight)
am          Transmission type (0 = automatic, 1 = manual)  
gear        Number of forward gears  
carb        Number of carburetors  
~~~~~

The first 10 rows of the dataset to get a overall view of the data.

```{r echo=FALSE}
head(mtcars)
```


## Appendic B - MPG vs Transmisstion Type Plot

```{r}
boxplot(mpg ~ am, data=mtcars, xlab="Transmission Type",
ylab="Miles per Gallon",
main="Automatic versus Manual Transmission MPG")
```