---
title: "Regression Models Exercises - Week 3"
author: "Shawn Rich"
date: "January 18, 2016"
output: html_document
---

```{r echo=FALSE, results="hide", warning=FALSE, message=FALSE}
library(ggplot2)
library(UsingR)
library(datasets)
library(dplyr)
data("galton")
data("father.son")
data("mtcars")
data("Seatbelts")
```


## Multivariable regression analysis

**1. Load the dataset Seatbelts as part of the datasets package via data(Seatbelts). Use as.data.frame to convert the object to a dataframe. Fit a linear model of driver deaths with kms and PetrolPrice as predictors. Interpret your results.**

```{r}
data("Seatbelts")
seatbelts <- as.data.frame(Seatbelts)
#head(seatbelts)
# normalize PetroPrice and mean center kms
seatbelts <- mutate(seatbelts,
                    pp = (PetrolPrice - mean(PetrolPrice)) / sd(PetrolPrice),
                    mm = kms/ 1000,
                    mmc = mm - mean(mm))

fit <- lm(DriversKilled ~ mmc + pp, data = seatbelts)
summary(fit)$coef
```


**2. Predict the number of driver deaths at the average kms and petrol levels.**

```{r}
data("Seatbelts")
seatbelts <- as.data.frame(Seatbelts)
head(seatbelts)
fit <- lm(DriversKilled ~ kms + PetrolPrice, data = seatbelts)
predict(fit, newdata = data.frame(kms = mean(seatbelts$kms), 
                                             PetrolPrice = mean(seatbelts$PetrolPrice)))
```


**3. Take the residual for DriversKilled having regressed out kms and an intercept and the residual for PetrolPrice having regressed out kms and an intercept. Fit a regression through the origin of the two residuals and show that it is the same as your coefficient obtained in question 1.**

```{r}
data("Seatbelts")
seatbelts <- as.data.frame(Seatbelts)
dk <- seatbelts$DriversKilled
kms <- seatbelts$kms
pp <- seatbelts$PetrolPrice
fitFull <- lm(dk ~ kms + pp)
# regress out kms from drivers killed and get residuals
edk <- resid(lm(dk ~ kms))
# regress out kms from petrol price killed and get residuals
epp <- resid(lm(pp ~ kms))

summary(lm(edk ~ epp -1))$coef
summary(fitFull)$coef
```


**4. Take the residual for DriversKilled having regressed out PetrolPrice and an intercept. Take the residual for kms having regressed out PetrolPrice and an intercept. Fit a regression through the origin of the two residuals and show that it is the same as your coefficient obtained in question 1.**

```{r}
# regress Petrol Price from Drivers Killed and kms 
# and compare this to fitFull summary
edk <- resid(lm(dk ~ pp))
ekms <- resid(lm(kms ~ pp))
summary(lm(edk ~ ekms - 1))$coef
summary(fitFull)$coef
```


## Multivariable examples and tricks

**1. Do exercise 1 of the previous chapter if you have not already. Load the dataset Seatbelts as part of the datasets package via data(Seatbelts). Use as.data.frame to convert the object to a dataframe. Fit a linear model of driver deaths with kms and PetrolPrice as predictors. Interpret your results.**

```{r}
data("Seatbelts")
seatbelts <- as.data.frame(Seatbelts)
#head(seatbelts)
# normalize PetroPrice and mean center kms
seatbelts <- mutate(seatbelts,
                    pp = (PetrolPrice - mean(PetrolPrice)) / sd(PetrolPrice),
                    mm = kms/ 1000,
                    mmc = mm - mean(mm))

fit <- lm(DriversKilled ~ pp + mmc, data = seatbelts)
summary(fit)$coef
```

**2. Repeat question 1 for the outcome being the log of the count of driver deaths. Interpret your coefficients. **

```{r}
data("Seatbelts")
seatbelts <- as.data.frame(Seatbelts)
#head(seatbelts)
# normalize PetroPrice and mean center kms
seatbelts <- mutate(seatbelts,
                    pp = (PetrolPrice - mean(PetrolPrice)) / sd(PetrolPrice),
                    mm = kms/ 1000,
                    mmc = mm - mean(mm))

fit <- lm(I(log(DriversKilled)) ~ pp + mmc, data = seatbelts)
summary(fit)$coef
```

**3. Refer to question 1. Add the dummy variable law and interpret the results. Repeat this question with a factor variable that you create called lawFactor that takes the levels No and Yes. Change the reference level from No to Yes.**

```{r}
data("Seatbelts")
seatbelts <- as.data.frame(Seatbelts)
#head(seatbelts)
# normalize PetroPrice and mean center kms
seatbelts <- mutate(seatbelts,
                    pp = (PetrolPrice - mean(PetrolPrice)) / sd(PetrolPrice),
                    mm = kms/ 1000,
                    mmc = mm - mean(mm))

fit <- lm(DriversKilled ~ pp + mmc + law, data = seatbelts)
summary(fit)$coef
fit <- lm(DriversKilled ~ pp + mmc + I(factor(law)), data = seatbelts)
summary(fit)$coef
```

**4. Discretize the PetrolPrice variable into four factor levels. Fit the linear model with this factor to see how R treats multiple level factor variables. **

```{r}
data("Seatbelts")
seatbelts <- as.data.frame(Seatbelts)
#head(seatbelts)
# normalize PetroPrice and mean center kms
seatbelts <- mutate(seatbelts,
                    pp = (PetrolPrice - mean(PetrolPrice)) / sd(PetrolPrice),
                    ppf = as.factor((pp<=-1.5)+(pp<=0)+(pp<=1.5)+(pp<Inf)),
                    mm = kms/ 1000,
                    mmc = mm - mean(mm))
fit <- lm(DriversKilled ~ ppf + kms + law, data=seatbelts)
summary(fit)$coef
```

**5. Perform the plot requested at the end of the last chapter.**

```{r}
```


## Adjustment

**1. Load the dataset Seatbelts as part of the datasets package via data(Seatbelts). Use as.data.frame to convert the object to a dataframe. Fit a linear model of driver deaths with kms and PetrolPrice as predictors. Interpret your results.**

```{r}
data("Seatbelts")
seatbelts <- as.data.frame(Seatbelts)
# normalize PetroPrice and mean center kms
seatbelts <- mutate(seatbelts,
                    pp = (PetrolPrice - mean(PetrolPrice)) / sd(PetrolPrice),
                    mm = kms/ 1000,
                    mmc = mm - mean(mm))
fit <- lm(DriversKilled ~ pp + mmc, data=seatbelts)
summary(fit)$coef
```

**2. Compare the kms coefficient with and without the inclusion of the PetrolPrice variable in the model.**

```{r}
fit0 <- lm(DriversKilled ~ mmc, data=seatbelts)
fit1 <- lm(DriversKilled ~ mmc + pp, data=seatbelts)
cor(seatbelts$pp,seatbelts$mmc)
summary(fit0)$coef
summary(fit1)$coef
# additional look at anova of the fits
anova(fit0, fit1)
```

**3. Compare the PetrolPrice coefficient with and without the inclusion fo the kms variable in the model.**

```{r}
fit0 <- lm(DriversKilled ~ pp, data=seatbelts)
fit1 <- lm(DriversKilled ~ pp + mmc, data=seatbelts)
cor(seatbelts$pp,seatbelts$mmc)
summary(fit0)$coef
summary(fit1)$coef
# additional look at anova of the fits
anova(fit0, fit1)
```


## Residuals, variation, diagnostics

**1. Load the dataset Seatbelts as part of the datasets package via data(Seatbelts). Use as.data.frame to convert the object to a dataframe. Fit a linear model of driver deaths with kms, PetrolPrice and law as predictors.**

```{r}
data("Seatbelts")
seatbelts <- as.data.frame(Seatbelts)
#head(seatbelts)
# normalize PetroPrice and mean center kms
seatbelts <- mutate(seatbelts,
                    pp = (PetrolPrice - mean(PetrolPrice)) / sd(PetrolPrice),
                    mm = kms/ 1000,
                    mmc = mm - mean(mm))

fit <- lm(DriversKilled ~ mmc + pp + law, data = seatbelts)
```

**2. Refer to question 1. Directly estimate the residual variation via the function resid. Compare with R’s residual variance estimate.**

```{r}
sum(resid(fit)^2) / (nrow(seatbelts) - 4)
summary(fit)$sigma^2
# is variance same with original kms and PetrolPrice?
fit2 <- lm(DriversKilled ~ kms + PetrolPrice + law, data = seatbelts)
summary(fit)$sigma^2
```

**3. Refer to question 1. Perform an analysis of diagnostic measures including, dffits, dfbetas, influence and hat diagonals.**

```{r}


#plot(dffits(fit))

#plot(dfbetas(fit))
#plot(influence(fit))
#hatvalues(fit)
```


## Multiple variables and model selection

**1. Load the dataset Seatbelts as part of the datasets package via data(Seatbelts). Use as.data.frame to convert the object to a dataframe. Fit a linear model of driver deaths with kms, PetrolPrice and law as predictors.**

```{r}
```

**2. Perform a model selection exercise to arrive at a final model.**

```{r}
data("Seatbelts")
seatbelts <- as.data.frame(Seatbelts)
#head(seatbelts)
# normalize PetroPrice and mean center kms
seatbelts <- mutate(seatbelts,
                    pp = (PetrolPrice - mean(PetrolPrice)) / sd(PetrolPrice),
                    mm = kms/ 1000,
                    mmc = mm - mean(mm))

fit <- lm(DriversKilled ~ mmc + pp + law, data = seatbelts)
```

