---
title: "Regression Models Exercises - Week 4"
author: "Shawn Rich"
date: "January 18, 2016"
output: pdf_document
---

```{r echo=FALSE, results="hide", warning=FALSE, message=FALSE}
library(ggplot2)
library(UsingR)
library(datasets)
library(dplyr)
data("galton")
data("father.son")
data("mtcars")
data("Seatbelts")
```
## Generalized Linear Models

1. True or false, generalized linear models transform the observed outcome.

False. 

```{r}

```


2. True or false, the interpretation of the coefficients in a GLM are on the scale of the link function. (Discuss.)

True.

```{r}

```


3. True or false, the generalized linear model assumes an exponential family for the outcome. (Discuss.)

True.

```{r}

```


4. True or false, GLM estimates are obtained by maximizing the likelihood. (Discuss.)

```{r}

```


5. True or false, some GLM distributions impose restrictions on the relationship between the mean and the variance. (Discuss.)

```{r}

```


## Binary GLMs

1. Load the dataset Seatbelts as part of the datasets package via data(Seatbelts). Use as.data.frame to convert the object to a dataframe. Create a new outcome variable for whether or not greater than 119 drivers were killed that month. Fit a logistic regression GLM with this variable as the outcome and kms, PetrolPrice and law as predictors. Interpret your parameters. 

```{r}
data("Seatbelts")
seatbelts <- as.data.frame(Seatbelts)
seatbelts <- mutate(seatbelts,
                    pp = (PetrolPrice - mean(PetrolPrice)) / sd(PetrolPrice),
                    mm = kms/ 1000,
                    mmc = mm - mean(mm),
                    dkb = ifelse(DriversKilled > 119, 1, 0))

fit <- glm(dkb ~ mmc + pp + law, family = "binomial", data = seatbelts)
round(summary(fit)$coef, 3)
```

2. Fit a binomial model with DriversKilled as the outcome and drivers as the total count with kms , PetrolPrice and law as predictors, interpret your results.

```{r}
data("Seatbelts")
seatbelts <- as.data.frame(Seatbelts)
seatbelts <- mutate(seatbelts,
                    pp = (PetrolPrice - mean(PetrolPrice)) / sd(PetrolPrice),
                    mm = kms/ 1000,
                    mmc = mm - mean(mm))

fit <- glm(DriversKilled ~ drivers + mmc + pp + law, data = seatbelts)
round(summary(fit)$coef, 3)
```

3. Refer to Question 1. Use the anova function to compare models with just law, law and PetrolPrice and all three predictors. 

```{r}
data("Seatbelts")
seatbelts <- as.data.frame(Seatbelts)
seatbelts <- mutate(seatbelts,
                    pp = (PetrolPrice - mean(PetrolPrice)) / sd(PetrolPrice),
                    mm = kms/ 1000,
                    mmc = mm - mean(mm),
                    dkb = ifelse(DriversKilled > 119, 1, 0))

fit1 <- glm(dkb ~ law, family = "binomial", data = seatbelts)
fit2 <- glm(dkb ~ law + pp, family = "binomial", data = seatbelts)
fit3 <- glm(dkb ~ law + pp + mmc, family = "binomial", data = seatbelts)
anova(fit1, fit2, fit3)
```


## Count Data

1. Load the dataset Seatbelts as part of the datasets package via data(Seatbelts). Use as.data.frame to convert the object to a dataframe. Fit a Poisson regression GLM with UKDriversKilled as the outcome and kms, PetrolPrice and law as predictors. Interpret your results.

```{r}
data("Seatbelts")
seatbelts <- as.data.frame(Seatbelts)
seatbelts <- mutate(seatbelts,
                    pp = (PetrolPrice - mean(PetrolPrice)) / sd(PetrolPrice),
                    mm = kms/ 1000,
                    mmc = mm - mean(mm))
fit <- glm(DriversKilled ~ mmc + pp + law, family = "poisson", data = seatbelts)
summary(fit)
```


2. Refer to question 1. Fit a linear model with the log of drivers killed as the outcome. Interpret your results.

```{r}
data("Seatbelts")
seatbelts <- as.data.frame(Seatbelts)
seatbelts <- mutate(seatbelts,
                    pp = (PetrolPrice - mean(PetrolPrice)) / sd(PetrolPrice),
                    mm = kms/ 1000,
                    mmc = mm - mean(mm))
fit <- lm(I(log(DriversKilled)) ~ mmc + pp + law, data = seatbelts)
summary(fit)

```


3. Refer to question 1. Fit your Poisson log-linear model with drivers as a log offset (to consider the proportion of drivers killed of those killed or seriously injured.)

```{r}
data("Seatbelts")
seatbelts <- as.data.frame(Seatbelts)
seatbelts <- mutate(seatbelts,
                    pp = (PetrolPrice - mean(PetrolPrice)) / sd(PetrolPrice),
                    mm = kms/ 1000,
                    mmc = mm - mean(mm))
fit0 <- glm(DriversKilled ~  pp + mmc + law, family = "poisson", data = seatbelts)
fit <- glm(DriversKilled ~  pp + mmc + law, family = "poisson", data = seatbelts, offset = log(drivers))
summary(fit0)$coef
summary(fit)$coef
```

4. Refer to Question 1. Use the anova function to compare models with just law, law and PetrolPrice and all three predictors.

```{r}
data("Seatbelts")
seatbelts <- as.data.frame(Seatbelts)
seatbelts <- mutate(seatbelts,
                    pp = (PetrolPrice - mean(PetrolPrice)) / sd(PetrolPrice),
                    mm = kms/ 1000,
                    mmc = mm - mean(mm))
fit1 <- glm(DriversKilled ~ law, family = "poisson", data = seatbelts)
fit2 <- glm(DriversKilled ~ law + pp, family = "poisson", data = seatbelts)
fit3 <- glm(DriversKilled ~ law + pp + mmc, family = "poisson", data = seatbelts)
anova(fit1, fit2, fit3)
```