---
title: "Relationship of Variables and Miles Per Gallon"
author: "Shawn Rich"
date: "January 26, 2016"
output: pdf_document
---

```{r warning=FALSE, echo=FALSE, error=FALSE, message=FALSE}
# load libraries
library(UsingR)
data("mtcars")
```

## Summary

In this report I will use the mtcars dataset to examine the relationship among the variables in the data and miles per gallon (MPG) as outcome. Specifically I will address whether the an automatic or manual transmission results in better gas milage for the car. To this end I will quantify the MPG difference between automatic and manual transmissions.

## Data Exploration

Description of the variables in the mtcars dataset is in [table 1] of the Appendix.

```{r}
head(mtcars)
```

## Appendix

**Table 1:** MTCARS dataset

Variable   | Description
---------- | -------------  
mpg        | Miles/(US) gallon  
cyl        | Number of cylinders  
disp       | Displacement (cu.in.)  
hp         | Gross horsepower  
drat       | Rear axle ratio  
wt         | Weight (1000 lbs)  
qsec       | 1/4 mile time  
vs         | V/S  
am         | Transmission (0 = automatic, 1 = manual)  
gear       | Number of forward gears  
carb       | Number of carburetors  



**Figure 1:** A Graph of Something
