---
title: "R Examples"
output:
  pdf_document: default
  html_document: default
---

## Drop Columns from a Data Frame

Create example data frame:

```{r}
df <- data.frame( a = 1:10, b = 2:11, c = 3:12 )
df
```

Select those you want to keep:

```{r}
df1 <- subset(df, select = c(a,c))
df1
```

Remove those you don't want (notice the minus sign):

```{r}
df2 <- subset(df, select = -c(a,c))
df2
```

## Ording Example

```{r}
dd <- data.frame(b = factor(c("Hi", "Med", "Hi", "Low"), 
      levels = c("Low", "Med", "Hi"), ordered = TRUE),
      x = c("A", "D", "A", "C"), y = c(8, 3, 9, 9),
      z = c(1, 1, 1, 2))

dd
```

```{r}
dd[with(dd, order(-z, b)), ]
```

## Replace/Update Column Value Based on Another

```{r}
df <- data.frame(n = c(1:12),
                 m = rep(c("?","+","h","k","m","b"),2))
df
```


```{r}
# convert t to character to make easier to compare
df$m <- as.character(df$m)

# set m as indicated by factor f
df$n[which(df$m %in% c("?","+"))] <- df$n[which(df$m %in% c("?","+"))] * 1
df$n[which(df$m == "h")] <- df$n[which(df$m == "h")] * 100
df$n[which(df$m == "k")] <- df$n[which(df$m == "k")] * 1000
df$n[which(df$m == "m")] <- df$n[which(df$m == "m")] * 1000000
df$n[which(df$m == "b")] <- df$n[which(df$m == "b")] * 1000000000

df
```

## IN and NOT IN

The example data.

```{r}
df <- data.frame(n = c(1:12),
                 m = rep(c("a","b","c","z"),3))
df
```

Rows where m is "a" or "c"

```{r}
x <- df[df$m %in% c("a","c"),]
 
x  
```

Rows where m is NOT "a" or "c"

```{r}
x <- df[!(df$m %in% c("a","c")),]
 
x  
```

## UpShift and LowerShift

The example data.

```{r}
df <- data.frame(n = c(1:12),
                 m = rep(c("9","b","1","z","+","?"),2))
df
```

UpShift column m

```{r}
df$m <- toupper(df$m)
df
```

LowerShift column m

```{r}
df$m <- tolower(df$m)
df
```

## Divide all Values in Column by Constant

```{r}
df <- data.frame(a = c(1:10), b = seq(2,20,by=2))
df
```

Divide column by 2

```{r}
df$b <- df$b/2
df
```

## Working with NA values

```{r}
df <- data.frame(col1 = c(1, 2, 3, NA, 5), col2 = c('A', 'B', NA, 'C', 'D'),
             col3 = c(9, NaN, 8, 7, 6))
df
```

List rows in which any column (at least one column) has an NA

```{r}
df[rowSums(is.na(df))>0,]
```

Change "na" to NA; sometime data is marked as NA with character string "na"
and we want really NA. Alos, the "na" turns columns into Factor variables

```{r}
df <- data.frame(col1 = c(1,2,"na",4), col2 = c(5,"na",7,8))
```

First, change "na" to NA

```{r}
df[is.na(df),]
```

The "na" also turns the colums to Factor variables, so we need to convert to integer