---
title: "T Confidence Intervals"
author: "Shawn Rich"
date: "September 30, 2015"
output: html_document
---

## Gosset's *t* distribution

* Invented by William Gosset in 1908 under pseudonym "Student". Often referred to as Student's t distribution.
* Has thicker tails than normal distribution.
* Assumes underlying data are iid Gaussian such that  
    $\frac{\bar{X} - \mu}{S/\sqrt{n}}$
* Interval is:  
    $\bar{X} \pm t_n-1*S/\sqrt{n}$, where $t_n-1$ is relevant quantile.

## Resources on the web

* [https://en.wikipedia.org/wiki/Student's_t-test](https://en.wikipedia.org/wiki/Student's_t-test)  
* [https://en.wikipedia.org/wiki/Student's_t-distribution](https://en.wikipedia.org/wiki/Student's_t-distribution)  
* [https://en.wikibooks.org/wiki/LaTeX/Mathematics](https://en.wikibooks.org/wiki/LaTeX/Mathematics) Very useful when typing up these notes.

## One Group (Sample) Version

*Example 1*  
An outbreak of Samonella-related illnes was attributed to ice cram produced at a certain factory. Scientist meassured the level of Salmonella in 9 randomly sampled batches of ice cream. The levels (in MPN/g) were:  

0.593, 0.142, 0.329, 0.691, 0.231, 0.793, 0.519, 0.392, 0.418  

Is there a evidence atht the mean level of Salmonella in the ice cream is greater than 0.3 MPN/g?

$H_0: \mu = 0.3$   use mu=0.3 in R code  
$H_a: \mu > 0.3$   use alternative = "greater" in R code  
    
```{r}
x <- c(0.593, 0.142, 0.329, 0.691, 0.231, 0.793, 0.519, 0.392, 0.418)
t.test(x, alternative = "greater", mu=0.3, conf.level = 0.95)
```

With p-value = 0.02927, there is moderately strong evidence that the mean Salmonella level in the ice cream is above 0.3 MPN/g. Reject null hypothesis in favor of the alternative.  

The same again, but using $\mu$ and $\sigma$ of the data. 

```{r}
x <- c(0.593, 0.142, 0.329, 0.691, 0.231, 0.793, 0.519, 0.392, 0.418)
mu <- mean(x)
n <- length(x)
df <- n - 1
s <- sd(x)
# 95% confidence interval:
mu + qt(0.95, df, lower.tail = FALSE)*s/sqrt(n)
```

The 95% confidence is same as the t.test result.

## Two Group (Sample) Version  

**Example 1 (independent samples)**  
6 subjects were given a drug treatment (treatment group) and an additional 6 others were given a placebo (control group). Perform a two-sample t-test for comparing the means of the treatment and control groups.

Let $\mu_1$ be the mean of the treatment group and $\mu_2$ the mean of the control group. We then have hypothesis of interest:

$H_0: \mu_1 - \mu_2 = 0$   use mu=0 in R code  
$H_a: \mu_1 - \mu_2 < 0$   use alternative = "less" in R code  

Assumming variances equal (var.equal = TRUE in R code).  
```{r}
Control <- c(91, 87, 99, 77, 88, 91)
Treat <- c(101, 110, 103, 93, 99, 104)
t.test(Control, Treat, alternative = "less", var.equal = TRUE)
```

Assumming variances not equal (var.equal = FALSE in R code).  
```{r}
Control <- c(91, 87, 99, 77, 88, 91)
Treat <- c(101, 110, 103, 93, 99, 104)
t.test(Control, Treat, alternative = "less", var.equal = FALSE)
```

Both the pooled t-test and the Welsh t-test give roughly the same results. Reject null hypothesis.

**Example 2 (dependent samples)**  
A study was conducted to test whether cars get better gas mileage on premium gas over regular. Each of 10 cars was first filled with either regular or premium gas, decided on a coin toss, and mileage recorded. Then mileage was recorded again with cars getting opposite gas than before. $\mu_1$ is mean of permium and $\mu_2$ is regular. Hypothesis:

$H_0: \mu_1 - \mu_2 = 0$   use mu=0 in R code (default)  
$H_a: \mu_1 - \mu_2 > 0$   use alternative = "greater" in R code  

Since these are dependent groups (samples), use paired = TRUE in R code.

```{r}
reg <- c(16, 20, 21, 22, 23, 22, 27, 25, 27, 28)
prem <- c(19, 22, 24, 24, 25, 25, 26, 26, 28, 32)
t.test(prem,reg,alternative = "greater", paired = TRUE)
```

Based on t statistic of 4.4721 and p-value of 0.00075 (very low), we reject null hypothesis. Strong evidence of a mean increase in mileage between regular and premium gasoline.

Now for the "manual" approach.

```{r}
reg <- c(16, 20, 21, 22, 23, 22, 27, 25, 27, 28)
prem <- c(19, 22, 24, 24, 25, 25, 26, 26, 28, 32)
difference <- prem - reg
mn <- mean(difference)
s <- sd(difference)
n <- 10
df <- n -1
# confidence interval
mn + c(-1,1)*qt(0.95,n-1)*s/sqrt(n)
```

**Example 3 (different sample sizes)**

*need something for this one.*


**A Tutorial**

Finish this tutorial here: [ww.ruf.rice.edu/~bioslabs/Stats_tutorial/ttest1.html](ww.ruf.rice.edu/~bioslabs/Stats_tutorial/ttest1.html)
```{r}
quercus_rubra <- c(6.4, 7.4, 10.4, 8.5, 9.2, 9.5, 11.8, 11.8, 9.3, 12.5, 13.7, 17)
quercus_alba <- c(11.4, 9.5, 5.8, 9.2, 11.8, 7.9, 7.9, 7.8, 5.3, 7.5, 7.1, 6.6)
mqr <- mean(quercus_rubra)
mqa <- mean(quercus_alba)
sqr <- sd(quercus_rubra)
sqa <- sd(quercus_alba)
```